---
title: "Training Log"
output:
  html_document: default
  pdf_document: default
theme: cerulean
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Last 7 Days
```{r echo=FALSE, message=FALSE}
library(reticulate)
library(dplyr)
library(ggplot2)
library(tidyr)
library(graphics)
library(gridExtra)
library(knitr)
library(plotly)
#options(scipen=999)
source_python("/Users/jslomas/Documents/JohnnyLomas.github.io/garmin_utilities.py")
# tryCatch(
#     expr = {
#         appendToRecord()
#     },
#     error = function(e){
#         print("Error while appending to record...")
#         print(e)
#     },
#     warning = function(w){
#     },
#     finally = {
#     }
# )


today <- Sys.Date()
days <- 7
now <- as.numeric(Sys.time())*1000
weekAgo <- now - (days)*24*60*60*1000
threeMonthsAgo <- now - (90)*24*60*60*1000

activities <- read.csv("/Users/jslomas/Documents/JohnnyLomas.github.io/periodOfRecord.csv")
activities$Type <- as.character(activities$Type)
activities[activities == "cross_country_skiing_ws"] <- "Backcountry_Skiing"
#activities[activities == "other"] <- "Backcountry_Skiing"
activities[activities == "treadmill_running"] <- "running"
activities$Type <- as.factor(activities$Type)

thisWeek <- filter(activities, Start_Time_Milli > weekAgo)

dateWhile <- 1
dateLevels <- c(format(today))
while(dateWhile < days){
  dateLevels <- c(format(today - dateWhile), dateLevels)
  dateWhile <- dateWhile + 1
}
thisWeek <- thisWeek %>% mutate(Date = format(as.Date(paste(Month, Day, Year, sep="/"), format = '%m/%d/%Y')))
thisWeek$Date <- factor(thisWeek$Date, levels=dateLevels)
totals <- getTotals(thisWeek, "all")
```

| Distance (mi)          | Vertical (ft)               | Time (hrs)              | Zone 0 (%)            | Zone 1 (%)            | Zone 2 (%)            | Zone 3 (%)            | Zone 4 (%)            | Zone 5 (%)            |
|---------------------------|-------------------------------|---------------------------|-----------------------|-----------------------|-----------------------|-----------------------|-----------------------|-----------------------|
| `r round(totals$Distance_Miles, digits=2)` | `r round(totals$Vertical_FT, digits=0)` | `r round(totals$Time_Hours, digits=2)` | `r round(totals$Z0_Percent, digits=2)` | `r round(totals$Z1_Percent, digits=2)` | `r round(totals$Z2_Percent, digits=2)` | `r round(totals$Z3_Percent, digits=2)` | `r round(totals$Z4_Percent, digits=2)` | `r round(totals$Z5_Percent, digits=2)` |

---


```{r echo=FALSE, message=FALSE, fig.width = 14, fig.height=6}
duration <- ggplot(data = thisWeek, aes(x=Date, y=Duration_Hours, fill=Type)) + 
  geom_col() +
  scale_x_discrete(drop = FALSE) +
  ylab("Time (hrs)") +
  xlab("") + 
  ggtitle("Duration") +
  theme(axis.text.x = element_text(angle = -45, hjust=0.2, vjust=0.2),
        legend.key.height = unit(0.5, 'cm'),
        legend.key.width = unit(0.2, 'cm'),
        legend.title = element_text(size=8), 
        legend.text = element_text(size=6))

zones <- ggplot(data = thisWeek %>% 
              select(Date, HR_Z1, HR_Z2, HR_Z3, HR_Z4, HR_Z5) %>%  
              pivot_longer(!c(Date), names_to="Zone", values_to="Time") %>% 
              group_by(Date, Zone) %>% 
              summarise(Time = sum(Time)), 
       aes(x=Date, y=Time, fill=Zone)) + 
  geom_col() +
  scale_x_discrete(drop = FALSE) +
  ylab("Time (hrs)") +
  xlab("") + 
  ggtitle("Heart Rate Zones") +
  theme(axis.text.x = element_text(angle = -45, hjust=0.2, vjust=0.2),
        legend.key.height = unit(0.5, 'cm'),
        legend.key.width = unit(0.2, 'cm'),
        legend.title = element_text(size=8), 
        legend.text = element_text(size=6))

vert <- ggplot(data = thisWeek %>% mutate(
      Vertical_Gain_Feet = Vertical_Gain_Feet/1000
    ), aes(x=Date, y=Vertical_Gain_Feet, fill=Type)) + 
  geom_col() +
  scale_x_discrete(drop = FALSE) +
  ylab("Vertical Feet (K)") +
  xlab("") +
  ggtitle("Vertical") +
  theme(axis.text.x = element_text(angle = -45, hjust=0.2, vjust=0.2),
        legend.key.height = unit(0.5, 'cm'),
        legend.key.width = unit(0.2, 'cm'),
        legend.title = element_text(size=8), 
        legend.text = element_text(size=6))

distance <- ggplot(data = thisWeek, aes(x=Date, y=Distance_Miles, fill=Type)) + 
  geom_col() +
  scale_x_discrete(drop = FALSE) +
  ylab("Distance (mi)") +
  xlab("") + 
  ggtitle("Distance") +
  theme(axis.text.x = element_text(angle = -45, hjust=0.2, vjust=0.2),
        legend.key.height = unit(0.5, 'cm'),
        legend.key.width = unit(0.2, 'cm'),
        legend.title = element_text(size=8), 
        legend.text = element_text(size=6))

durationGrob <- ggplotGrob(duration)
zonesGrob <- ggplotGrob(zones)
vertGrob <- ggplotGrob(vert)
distanceGrob <- ggplotGrob(distance)

grid.arrange(durationGrob, zonesGrob, vertGrob, distanceGrob)
```

---

```{r echo=FALSE}
kable(thisWeek %>% select(DayofWeek,Type,Distance_Miles,Duration_Hours,Vertical_Gain_Feet,Rating,Notes) %>% mutate_if(is.numeric, round, 2) %>% mutate(Vertical_Gain_Feet = round(Vertical_Gain_Feet, digits=0)),
      caption = "This week's training log.")
```

## 3 Month Overview
Totals are aggregated on Monday of each week. 
```{r echo=FALSE, message=FALSE, fig.height=8}
#activities <- getActivitiesByPeriod(90, "weekly", FALSE)
firstDayofWeek <- function(milli){
  dayofweek <- as.numeric(strftime(as.POSIXct((milli) / 1000, origin = "1970-01-01"), format = "%u"))
  milli1 <- milli - ((dayofweek-1)*24*60*60*1000)
  return(format(as.POSIXct(milli1 / 1000, origin = "1970-01-01"), "%D"))
}

getWeeklies <- function(table, endDateMilli){
  weekOfYear <- strftime(as.POSIXct(endDateMilli / 1000, origin = "1970-01-01"), format = "%V")
  year <- strftime(as.POSIXct(endDateMilli / 1000, origin = "1970-01-01"), format = "%Y")
  month <- as.numeric(strftime(as.POSIXct(endDateMilli / 1000, origin = "1970-01-01"), format = "%m"))
  
  weeklies <- table %>% 
              filter(Year >= year & 
                     Month >= month & 
                     Week >= weekOfYear) %>%
              group_by(Year) %>% 
              group_by(Week) %>%
              summarise(Start_Time_Milli = min(Start_Time_Milli),
                     Distance_Miles = sum(Distance_Miles),
                     Duration_Hours = sum(Duration_Hours),
                     Vertical_Gain_Feet = sum(Vertical_Gain_Feet),
                     HR_Z1 = sum(HR_Z1),
                     HR_Z2 = sum(HR_Z2),
                     HR_Z3 = sum(HR_Z3),
                     HR_Z4 = sum(HR_Z4),
                     HR_Z5 = sum(HR_Z5),
                     Day = min(Day),
                     Year = min(Year),
                     Month = min(Month))
}
weeklySummary <- getWeeklies(activities, threeMonthsAgo)

weeklySummary <- weeklySummary %>% mutate(date = firstDayofWeek(Start_Time_Milli))
duration <- ggplot(data = weeklySummary, aes(x=date, y=Duration_Hours)) + 
  geom_col() +
  scale_x_discrete(drop = FALSE) +
  ylab("Time (hrs)") +
  xlab("")+
  ggtitle("Duration") +
  theme(axis.text.x = element_text(angle = -45, hjust=0.2, vjust=0.2))

#zones <- ggplot(data = activities %>% 
 #             select(DayofWeek, HR_Z1, HR_Z2, HR_Z3, HR_Z4, HR_Z5) %>%  
#              pivot_longer(!c(DayofWeek), names_to="Zone", values_to="Time") %>% 
 #             group_by(DayofWeek, Zone) %>% 
  #            summarise(Time = sum(Time)), 
   #    aes(x=DayofWeek, y=Time, fill=Zone)) + 
  #geom_col() +
  #scale_x_discrete(drop = FALSE) +
  #ylab("Time (hrs)") +
  #xlab("") + 
  #ggtitle("Heart Rate Zones") +
  #theme(axis.text.x = element_text(angle = -45, hjust=0.2, vjust=0.2),
   #     legend.key.height = unit(0.5, 'cm'),
    #    legend.key.width = unit(0.2, 'cm'),
     #   legend.title = element_text(size=8), 
      #  legend.text = element_text(size=6))

vert <- ggplot(data = weeklySummary %>% mutate(
      Vertical_Gain_Feet = Vertical_Gain_Feet/1000
    ), aes(x=date, y=Vertical_Gain_Feet)) + 
  geom_col() +
  scale_x_discrete(drop = FALSE) +
  ylab("Vertical Feet (Thousands)") +
  xlab("") +
  ggtitle("Vertical") +
  theme(axis.text.x = element_text(angle = -45, hjust=0.2, vjust=0.2))

distance <- ggplot(data = weeklySummary, aes(x=date, y=Distance_Miles)) + 
  geom_col() +
  scale_x_discrete(drop = FALSE) +
  ylab("Distance (mi)") +
  xlab("") + 
  ggtitle("Distance") +
  theme(axis.text.x = element_text(angle = -45, hjust=0.2, vjust=0.2))

durationGrob <- ggplotGrob(duration)
#zonesGrob <- ggplotGrob(zones)
vertGrob <- ggplotGrob(vert)
distanceGrob <- ggplotGrob(distance)

grid.arrange(durationGrob, vertGrob, distanceGrob)
```

## Yearly Cumulatives 
```{r message=FALSE,echo=FALSE}
#Cumulatives
cumulative <- activities %>% 
           arrange(Start_Time_Milli) %>%
           mutate(DayofYear = as.numeric(strftime(as.POSIXct(Start_Time_Milli / 1000, origin = "1970-01-01"), format = "%j"))) %>%
           group_by(Year) %>%
           mutate(Vertical_Gain_Feet = Vertical_Gain_Feet/1000) %>%
           mutate(CumVert = cumsum(Vertical_Gain_Feet)) %>%
           mutate(CumHours = cumsum(Duration_Hours)) %>%
           mutate(CumDist = cumsum(Distance_Miles))

cumulative$Year <-factor(cumulative$Year)

cumVert <- ggplot(data = cumulative, aes(x = DayofYear, y = CumVert, group =Year, color = Year)) + 
  geom_line() + 
  theme(axis.text.x = element_text(angle=0, hjust = 1)) +
  ylab("Cumulative Vertical (ft*k)") +
  xlab("Julian Day") +
  ggtitle("Vertical (All Sports)") #+ scale_x_discrete(labels = running$Month)
ggplotly(cumVert)

cumHours <- ggplot(data = cumulative, aes(x = DayofYear, y = CumHours, group =Year, color = Year)) + 
  geom_line() + 
  theme(axis.text.x = element_text(angle=0, hjust = 1)) +
  ylab("Cumulative Duration (hrs)") +
  xlab("Julian Day") +
  ggtitle("Duration (All Sports)") #+ scale_x_discrete(labels = running$Month)
ggplotly(cumHours)

cumDist <- ggplot(data = cumulative, aes(x = DayofYear, y = CumDist, group =Year, color = Year)) + 
  geom_line() + 
  theme(axis.text.x = element_text(angle=0, hjust = 1)) +
  ylab("Cumulative Distance (mi)") +
  xlab("Julian Day") +
  ggtitle("Distance (All Sports)") #+ scale_x_discrete(labels = running$Month)
ggplotly(cumDist)
```

## By Sport

```{r message=FALSE,echo=FALSE}
cumulative_types <- activities %>%
           filter(Type == "Backcountry_Skiing" | Type == "strength_training" | Type == "running" | Type == "mountain_biking" | Type == "cycling" | Type == "hiking" | Type == "other") %>%
           arrange(Start_Time_Milli) %>%
           mutate(DayofYear = as.numeric(strftime(as.POSIXct(Start_Time_Milli / 1000, origin = "1970-01-01"), format = "%j"))) %>%
           group_by(Year, Type) %>%
           mutate(Vertical_Gain_Feet = Vertical_Gain_Feet/1000) %>%
           mutate(CumVert = cumsum(Vertical_Gain_Feet)) %>%
           mutate(CumHours = cumsum(Duration_Hours)) %>%
           mutate(CumDist = cumsum(Distance_Miles))

cumulative_types$Type <- as.character(cumulative_types$Type)
cumulative_types$Type <- as.factor(cumulative_types$Type)
cumulative_types$Year <-factor(cumulative_types$Year)

ggplot(data = cumulative_types, aes(x = DayofYear, y = CumVert, group =Year, color = Year)) + 
  geom_line() + 
  theme(axis.text.x = element_text(angle=0, hjust = 1)) +
  ylab("Cumulative Vertical (ft*k)") +
  xlab("Julian Day") +
  ggtitle("Vertical") +
  facet_wrap(~Type)

ggplot(data = cumulative_types, aes(x = DayofYear, y = CumHours, group =Year, color = Year)) + 
  geom_line() + 
  theme(axis.text.x = element_text(angle=0, hjust = 1)) +
  ylab("Cumulative Duration (hrs)") +
  xlab("Julian Day") +
  ggtitle("Duration") +
  facet_wrap(~Type)

ggplot(data = cumulative_types, aes(x = DayofYear, y = CumDist, group =Year, color = Year)) + 
  geom_line() + 
  theme(axis.text.x = element_text(angle=0, hjust = 1)) +
  ylab("Cumulative Distance (mi)") +
  xlab("Julian Day") +
  ggtitle("Distance") +
  facet_wrap(~Type)
```


