---
title: "Training Log"
output:
  html_document: default
  pdf_document: default
theme: cerulean
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Last 7 Days
```{r echo=FALSE, message=FALSE}
library(reticulate)
library(dplyr)
use_python("/usr/bin/python3")
source_python("/Users/jslomas/Documents/JohnnyLomas.github.io/garmin_utilities.py")
days <- 7
activities <- getActivitiesByPeriod(days)

#activities$DayofWeek <- factor(activities$DayofWeek, levels=c("Monday","Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"))

#activities <- activities %>% 
 # arrange(desc(Start_Time_Milli)) %>%
  #mutate(Date = as.Date(paste(Month, Day, Year, sep="/"), format = '%m/%d/%Y'))
dateWhile <- 1
today <- Sys.Date()
dateLevels <- c(format(today))
while(dateWhile <= days){
  dateLevels <- c(format(today - dateWhile), dateLevels)
  dateWhile <- dateWhile + 1
}
activities <- activities %>% mutate(Date = format(as.Date(paste(Month, Day, Year, sep="/"), format = '%m/%d/%Y')))
activities$Date <- factor(activities$Date, levels=dateLevels)
totals <- getTotals(activities, "all")
```

| Distance (mi)          | Vertical (ft)               | Time (hrs)              | Zone 0 (%)            | Zone 1 (%)            | Zone 2 (%)            | Zone 3 (%)            | Zone 4 (%)            | Zone 5 (%)            |
|---------------------------|-------------------------------|---------------------------|-----------------------|-----------------------|-----------------------|-----------------------|-----------------------|-----------------------|
| `r round(totals$Distance_Miles, digits=2)` | `r round(totals$Vertical_FT, digits=0)` | `r round(totals$Time_Hours, digits=2)` | `r round(totals$Z0_Percent, digits=2)` | `r round(totals$Z1_Percent, digits=2)` | `r round(totals$Z2_Percent, digits=2)` | `r round(totals$Z3_Percent, digits=2)` | `r round(totals$Z4_Percent, digits=2)` | `r round(totals$Z5_Percent, digits=2)` |

---


```{r echo=FALSE, message=FALSE, fig.width = 14, fig.height=6}
library(ggplot2)
library(tidyr)
library(graphics)
library(gridExtra)

duration <- ggplot(data = activities, aes(x=Date, y=Duration_Hours, fill=Type)) + 
  geom_col() +
  scale_x_discrete(drop = FALSE) +
  ylab("Time (hrs)") +
  xlab("") + 
  ggtitle("Duration") +
  theme(axis.text.x = element_text(angle = -45, hjust=0.2, vjust=0.2),
        legend.key.height = unit(0.5, 'cm'),
        legend.key.width = unit(0.2, 'cm'),
        legend.title = element_text(size=8), 
        legend.text = element_text(size=6))

zones <- ggplot(data = activities %>% 
              select(Date, HR_Z1, HR_Z2, HR_Z3, HR_Z4, HR_Z5) %>%  
              pivot_longer(!c(Date), names_to="Zone", values_to="Time") %>% 
              group_by(Date, Zone) %>% 
              summarise(Time = sum(Time)), 
       aes(x=Date, y=Time, fill=Zone)) + 
  geom_col() +
  scale_x_discrete(drop = FALSE) +
  ylab("Time (hrs)") +
  xlab("") + 
  ggtitle("Heart Rate Zones") +
  theme(axis.text.x = element_text(angle = -45, hjust=0.2, vjust=0.2),
        legend.key.height = unit(0.5, 'cm'),
        legend.key.width = unit(0.2, 'cm'),
        legend.title = element_text(size=8), 
        legend.text = element_text(size=6))

vert <- ggplot(data = activities %>% mutate(
      Vertical_Gain_Feet = Vertical_Gain_Feet/1000
    ), aes(x=Date, y=Vertical_Gain_Feet, fill=Type)) + 
  geom_col() +
  scale_x_discrete(drop = FALSE) +
  ylab("Vertical Feet (K)") +
  xlab("") +
  ggtitle("Vertical") +
  theme(axis.text.x = element_text(angle = -45, hjust=0.2, vjust=0.2),
        legend.key.height = unit(0.5, 'cm'),
        legend.key.width = unit(0.2, 'cm'),
        legend.title = element_text(size=8), 
        legend.text = element_text(size=6))

distance <- ggplot(data = activities, aes(x=Date, y=Distance_Miles, fill=Type)) + 
  geom_col() +
  scale_x_discrete(drop = FALSE) +
  ylab("Distance (mi)") +
  xlab("") + 
  ggtitle("Distance") +
  theme(axis.text.x = element_text(angle = -45, hjust=0.2, vjust=0.2),
        legend.key.height = unit(0.5, 'cm'),
        legend.key.width = unit(0.2, 'cm'),
        legend.title = element_text(size=8), 
        legend.text = element_text(size=6))

durationGrob <- ggplotGrob(duration)
zonesGrob <- ggplotGrob(zones)
vertGrob <- ggplotGrob(vert)
distanceGrob <- ggplotGrob(distance)

grid.arrange(durationGrob, zonesGrob, vertGrob, distanceGrob)
```

---

```{r echo=FALSE}
library(knitr)
kable(activities %>% select(DayofWeek,Type,Distance_Miles,Duration_Hours,Vertical_Gain_Feet,Rating,Notes) %>% mutate_if(is.numeric, round, 2) %>% mutate(Vertical_Gain_Feet = round(Vertical_Gain_Feet, digits=0)),
      caption = "This week's training log.")
```

## 3 Month Overview
Totals are aggregated on Monday of each week. 
```{r echo=FALSE, message=FALSE, fig.height=8}
activities <- getActivitiesByPeriod(90, "weekly", FALSE)
activities <- activities %>% mutate(date = format(as.POSIXct(Start_Time_Milli / 1000, origin = "1970-01-01"), "%D"))
duration <- ggplot(data = activities, aes(x=date, y=Duration_Hours)) + 
  geom_col() +
  scale_x_discrete(drop = FALSE) +
  ylab("Time (hrs)") +
  xlab("")+
  ggtitle("Duration") +
  theme(axis.text.x = element_text(angle = -45, hjust=0.2, vjust=0.2))

#zones <- ggplot(data = activities %>% 
 #             select(DayofWeek, HR_Z1, HR_Z2, HR_Z3, HR_Z4, HR_Z5) %>%  
#              pivot_longer(!c(DayofWeek), names_to="Zone", values_to="Time") %>% 
 #             group_by(DayofWeek, Zone) %>% 
  #            summarise(Time = sum(Time)), 
   #    aes(x=DayofWeek, y=Time, fill=Zone)) + 
  #geom_col() +
  #scale_x_discrete(drop = FALSE) +
  #ylab("Time (hrs)") +
  #xlab("") + 
  #ggtitle("Heart Rate Zones") +
  #theme(axis.text.x = element_text(angle = -45, hjust=0.2, vjust=0.2),
   #     legend.key.height = unit(0.5, 'cm'),
    #    legend.key.width = unit(0.2, 'cm'),
     #   legend.title = element_text(size=8), 
      #  legend.text = element_text(size=6))

vert <- ggplot(data = activities %>% mutate(
      Vertical_Gain_Feet = Vertical_Gain_Feet/1000
    ), aes(x=date, y=Vertical_Gain_Feet)) + 
  geom_col() +
  scale_x_discrete(drop = FALSE) +
  ylab("Vertical Feet (Thousands)") +
  xlab("") +
  ggtitle("Vertical") +
  theme(axis.text.x = element_text(angle = -45, hjust=0.2, vjust=0.2))

distance <- ggplot(data = activities, aes(x=date, y=Distance_Miles)) + 
  geom_col() +
  scale_x_discrete(drop = FALSE) +
  ylab("Distance (mi)") +
  xlab("") + 
  ggtitle("Distance") +
  theme(axis.text.x = element_text(angle = -45, hjust=0.2, vjust=0.2))

durationGrob <- ggplotGrob(duration)
#zonesGrob <- ggplotGrob(zones)
vertGrob <- ggplotGrob(vert)
distanceGrob <- ggplotGrob(distance)

grid.arrange(durationGrob, vertGrob, distanceGrob)
```

## Yearly Cumulatives 
```{r message=FALSE,echo=FALSE}
#TODO: Cumulative season pacing charts by activity
#ggplot(data = df, aes(x = 1:5, y = cumsum(val))) + geom_line() + 
       # geom_point() + theme(axis.text.x = element_text(angle=90, hjust = 1)) + 
        #scale_x_discrete(labels = df$date) + xlab("Date")

activities <- getActivitiesByPeriod(5000, "daily", FALSE)

#Cumulative running over the calendar year
running <- activities %>% 
           filter(Type == "running") %>% 
           arrange(Start_Time_Milli) %>%
           mutate(DayofYear = ((Month-1)*30)+Day) %>%
           group_by(Year) %>%
           mutate(CumDist = cumsum(Distance_Miles))

running$Year <-factor(running$Year)

ggplot(data = running, aes(x = DayofYear, y = CumDist, group =Year, color = Year)) + 
  geom_line() + 
  theme(axis.text.x = element_text(angle=0, hjust = 1)) +
  ylab("Cumulative Distance (mi)") +
  xlab("Julian Day") +
  ggtitle("Running Distance") #+ scale_x_discrete(labels = running$Month)

#Cumulative vertical gain over the calendar year
vertical <- activities %>% 
           arrange(Start_Time_Milli) %>%
           mutate(DayofYear = ((Month-1)*30)+Day) %>%
           group_by(Year) %>%
           mutate(Vertical_Gain_Feet = Vertical_Gain_Feet/1000) %>%
           mutate(CumVert = cumsum(Vertical_Gain_Feet))

vertical$Year <-factor(vertical$Year)

ggplot(data = vertical, aes(x = DayofYear, y = CumVert, group =Year, color = Year)) + 
  geom_line() + 
  theme(axis.text.x = element_text(angle=0, hjust = 1)) +
  ylab("Cumulative Vertical (ft*k)") +
  xlab("Julian Day") +
  ggtitle("Vertical (All Sports)") #+ scale_x_discrete(labels = running$Month)

#Cumulative duration over the calendar year
hours <- activities %>% 
           arrange(Start_Time_Milli) %>%
           mutate(DayofYear = ((Month-1)*30)+Day) %>%
           group_by(Year) %>%
           mutate(CumHours = cumsum(Duration_Hours))

hours$Year <- factor(hours$Year)

ggplot(data = hours, aes(x = DayofYear, y = CumHours, group =Year, color = Year)) + 
  geom_line() + 
  theme(axis.text.x = element_text(angle=0, hjust = 1)) +
  ylab("Cumulative Duration (hrs)") +
  xlab("Julian Day") +
  ggtitle("Duration (All Sports)") #+ scale_x_discrete(labels = running$Month)

#Cumulative distance over the calendar year
distance <- activities %>% 
           arrange(Start_Time_Milli) %>%
           mutate(DayofYear = ((Month-1)*30)+Day) %>%
           group_by(Year) %>%
           mutate(CumDist = cumsum(Distance_Miles))

distance$Year <- factor(distance$Year)

ggplot(data = distance, aes(x = DayofYear, y = CumDist, group =Year, color = Year)) + 
  geom_line() + 
  theme(axis.text.x = element_text(angle=0, hjust = 1)) +
  ylab("Cumulative Distance (mi)") +
  xlab("Julian Day") +
  ggtitle("Distance (All Sports)") #+ scale_x_discrete(labels = running$Month)
```

